---
title: "HW2"
author: "Christina Dai cd3347"
output: github_document
---

```{r setup}

library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

Loading the instacart data:

```{r}
library(p8105.datasets)
data("instacart")
```

This dataset contains information from the grocery delivery service Instacart. Customers can enlist a shopper, employed by Instacart, to do their grocery shopping for them at various grocery stores in their area. The shopper will then deliver the groceries to their door. There are `nrow(instacart)` rows and `ncol(instacart)` columns total. In this dataset, each observation corresponds to a specific product and includes information such as `product_name`, `aisle`, `department`, and `days_since_prior_order`. For example, the first row in the dataset is Bulgarian Yogurt in the dairy and eggs department in the yogurt aisle.

How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs))
```

There are 134 different aisles in the data and the most items are ordered from the fresh vegetables and fresh fruits aisles. These two aisles had nearly twice as many orders than the 3rd most ordered from aisle, which was packaged vegetables and fruits.

Make a plot for `aisles`:

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  filter(n_obs > 10000) %>% 
  arrange(desc(n_obs)) %>%
  ggplot(aes(x = aisle, y = n_obs)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”:

```{r}
aisles_pop = 
  instacart %>% 
    filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
    group_by(aisle, product_name) %>% 
    summarize(n_prod = n()) %>% 
    arrange(desc(n_prod)) %>% 
  group_by(aisle) %>% 
  slice(1:3) %>% 
  knitr::kable()
```

Table for hour of day Pink Lady Apples and Coffee Ice Cream are ordered for each day of the week:

```{r}
apples_coffee = 
  instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow, order_hour_of_day)
```


### Problem 2

Loading the BRFSS data:

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```


Data cleaning:

```{r}
brfss_clean = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(state = locationabbr, state_county = locationdesc) %>% 
  filter(topic == "Overall Health") %>% 
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>% 
  mutate(response = as.factor(response)) %>% 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
locations2002 = 
  brfss_clean %>% 
    group_by(state) %>% 
    filter(year == 2002) %>% 
    summarize(n_counties = n_distinct(state_county)) %>% 
    filter(n_counties >= 7)
```


```{r}
locations2010 = 
  brfss_clean %>% 
    group_by(state) %>% 
    filter(year == 2010) %>% 
    summarize(n_counties = n_distinct(state_county)) %>% 
    filter(n_counties >= 7)
```


In 2002, CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA were observed at 7 or more locations.

In 2010, CT, FL, MA, NC, NJ, and PA were observed at 7 or more locations.


Spaghetti plot:

Modifying the data

```{r}
excellent = 
  brfss_clean %>% 
  filter(response == "Excellent") %>% 
  group_by(year, state, state_county) %>% 
  summarize(data_value_avg = mean(data_value))
```










