---
title: "HW2"
author: "Christina Dai cd3347"
output: github_document
---

```{r setup}

library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

### Loading the instacart data:

```{r}
library(p8105.datasets)
data("instacart")
```

This dataset contains information from the grocery delivery service Instacart. Customers can enlist a shopper, employed by Instacart, to do their grocery shopping for them at various grocery stores in their area. The shopper will then deliver the groceries to their door. There are `nrow(instacart)` rows and `ncol(instacart)` columns total. In this dataset, each observation corresponds to a specific product and includes information such as `product_name`, `aisle`, `department`, and `days_since_prior_order`. For example, the first row in the dataset is Bulgarian Yogurt in the dairy and eggs department in the yogurt aisle.

### How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs))
```

There are 134 different aisles in the data and the most items are ordered from the fresh vegetables and fresh fruits aisles. These two aisles had nearly twice as many orders than the 3rd most ordered from aisle, which was packaged vegetables and fruits.

### Make a plot for `aisles`:

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  filter(n_obs > 10000) %>% 
  arrange(desc(n_obs)) %>%
  ggplot(aes(x = aisle, y = n_obs)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”:

```{r}
aisles_pop = 
  instacart %>% 
    filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
    group_by(aisle, product_name) %>% 
    summarize(n_prod = n()) %>% 
    arrange(desc(n_prod)) %>% 
  group_by(aisle) %>% 
  slice(1:3) 
  knitr::kable(aisles_pop)
```

### Table for hour of day Pink Lady Apples and Coffee Ice Cream are ordered for each day of the week:

```{r}
apples_coffee = 
  instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hr = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hr
  )
  knitr::kable(apples_coffee)
```


# Problem 2

### Loading the BRFSS data:

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```


### Data cleaning:

```{r}
brfss_clean = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(state = locationabbr, state_county = locationdesc) %>% 
  filter(topic == "Overall Health") %>% 
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>% 
  mutate(response = as.factor(response)) %>% 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

### In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
locations2002 = 
  brfss_clean %>% 
    group_by(state) %>% 
    filter(year == 2002) %>% 
    summarize(n_counties = n_distinct(state_county)) %>% 
    filter(n_counties >= 7)
```


```{r}
locations2010 = 
  brfss_clean %>% 
    group_by(state) %>% 
    filter(year == 2010) %>% 
    summarize(n_counties = n_distinct(state_county)) %>% 
    filter(n_counties >= 7)
```


In 2002, CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA were observed at 7 or more locations.

In 2010, CT, FL, MA, NC, NJ, and PA were observed at 7 or more locations.


### Spaghetti plot:

Modifying the data

```{r}
excellent = 
  brfss_clean %>% 
  filter(response == "Excellent") %>% 
  group_by(year, state) %>% 
  summarize(data_value_avg = mean(data_value))
```

Make the spaghetti plot:

```{r}
excellent %>% 
  ggplot(aes(x = year, y = data_value_avg, group = state, color = state)) +  
  geom_line() + 
  theme(legend.position = "none") + 
  labs(title = "Data Value Average by State from 2002-2010")
```

This plot shows the average of the data values within states for only "Excellent" responses. Most states have a data value average between around 17 to 29 for the years from 2002 to 2010. There are a couple outliers with a particularly low data value average, but otherwise, most states are consistent with each other and over time.


### Make the two-panel plot

For 2006:

```{r}
data_value2006 = 
  brfss_clean %>% 
    filter(year == 2006 & state == "NY") %>% 
    ggplot(aes(x = response, y = data_value)) + 
    geom_boxplot() + 
    labs(title = "Data Values for Response Levels 2006")
```

For 2010:

```{r}
data_value2010 = 
  brfss_clean %>% 
    filter(year == 2010 & state == "NY") %>% 
    ggplot(aes(x = response, y = data_value)) + 
    geom_boxplot() + 
    labs(title = "Data Values for Response Levels 2010")
```

Now patch them together:

```{r}
data_value2006 + data_value2010
```

From this plot, we can see that the data value for responses "Good" and "Very good" are higher than the others, with "Poor" at the lowest around a value of 3. Generally, it looks like a slightly left skewed distribution for both 2006 and 2010.

# Problem 3

### Load the datasets:

```{r}
nhanes_accel = 
  read_csv("./data/nhanes_accel.csv")
```

```{r}
nhanes_covar = 
  read_csv("./data/nhanes_covar.csv", skip = 4) %>% 
  drop_na()
```

### Merge the datasets and tidy data:

```{r}
nhanes_combined = 
  left_join(nhanes_covar, nhanes_accel, by = "SEQN") %>% 
  filter(age >= 21) %>% 
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) %>% 
  mutate(sex = as.factor(sex)) %>% 
  mutate(sex = factor(sex, levels = c("male", "female"))) %>% 
  mutate(education = recode(education, `1` = "<HS", `2` = "=HS", `3` = ">HS")) %>% 
  mutate(education = as.factor(education))
```

### Table and plot for men and women in each education category:

```{r}
sex_education = 
  nhanes_combined %>% 
  select(sex, education) %>% 
  group_by(sex, education) %>% 
  summarize(n_obs = n()) %>% 
  pivot_wider(
    names_from = education,
    values_from = n_obs
  )
  knitr::kable(sex_education)
```

Age distributions for men and women:

```{r}
sex_age = 
  nhanes_combined %>% 
  ggplot(aes(x = education, y = age, fill = sex)) + 
  geom_boxplot(alpha = 0.5) + 
  labs(title = "Distribution of Age by Sex and Education Level")
```


This table tells us the number of men and women in each education category. From the data, we can see that there are about the same number of men and women who have a less than HS education level and more than HS level. There are 12 more women who have an equal to HS education than men. Overall, more participants have a greater than HS education than a less than or equal to HS education.


### Aggregate activity plots:

```{r}
nhanes_combined %>% 
  mutate(total_act = rowSums(across(c(min1:min1440)))) %>% 
  select(sex, age, education, total_act) %>% 
  ggplot(aes(x = age, y = total_act, group = sex, color = sex)) + 
  geom_point(alpha = 0.5) + 
  facet_grid(. ~ education) + 
  geom_smooth(se = FALSE) + 
  labs(title = "Total Activity Level Based on Age and Education")
```

In this plot, we see how total activity level in minutse changes based on age and education level. For those with a less than HS education, younger people generally have a higher activity level than those with a equivalent to HS education and those with a greater than HS education. 
Activity levels trends based on age are similar between males and females with a less than HS education level. For those with a HS equivalent education level, females and males at age 21 have similar activity levels, but as participants get older, the females have higher activity levels until about age 60, when they again meet with the males' activity levels. For those with a greater than HS education, females have slightly higher activity levels than males across all ages.


### Activity over the course of a day plot:

```{r}
nhanes_combined %>% 
  pivot_longer(
    min1:min1440,
    names_to = "min_count",
    values_to = "minutes"
  ) %>% 
  ggplot(aes(x = min_count, y = minutes, group = sex, color = sex)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ education) + 
  labs(title = "Activity over 24 hrs Based on Sex and Education level",
       x = "24 hour period") + 
  theme(axis.text.x = element_blank())
```

This graph shows activity levels in minutes over the course of the 24 hour period participants were measured through by education level. In general, participants were more active for the first half of their 24-hour period, and then activity dipped briefly before going back up again for the remainder. Those with a less than HS education and an equivalent to HS education had similar 24-hour activity time courses, and males and females seemed to be pretty similar. For those with a more than HS education, there were more outliers with higher activity levels (more males earlier in the 24-hour perido and more females later in the 24-hour period). However, the aggregate curves shown seemed to be uniform with the other education levels.


